Proposed Solution

### 2017-01-25 CAV17/secs/trans.tex

Fig.~\ref{fig:ptypes} provides a fragment of a calculus for type judgments for deriving Boolean equality rules with probabilistic types.
We use the notation, $\ptype \models p$, to denote a probabilistic type whose counts hold on the predicate $p$. For a $\ptype$ with $|\tau_i|$ 
we denote $c_i$, and $*$ is shorthand for all basic types. When there is no evidence for a type $\tau$ we call this empty type ({\scriptsize EMPTY}). Initially, before we count any type, all types are set to empty. 
In the {\scriptsize BOOL\_RESOLVE} rule, a keyword with a probabilistic type $k : \ptype \in \tilde{\mathcal{T}}$ can be resolved to a basic type with enough evidence.
The definition of enough evidence must be empirically determined by the user depending on the quality of the learning set.

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=Empty]
{\\}
{k:[\ast = 0]}
\and
\inferrule* [Right=bool\_int]
{(k,v) \in C \\ v \in \{0,1\} \\ k:\ptype}
{k:\ptype\update[int\pluseq 1,bool\pluseq 1]}
\and
\inferrule* [Right=bool]
{(k,v) \in C \\ v \in \{ON,OFF\} \\ k:\ptype}
{k:\ptype\update[bool\pluseq 1]}
\\
\and
\inferrule* [Right=bool\_resolve]
{k:\ptype \\ \ptype \models |bool| \geq M \land |\ast| \leq N}
{k :: bool}
\\
\and
\inferrule* [Right=eq\_rule]
{k_1 :: \tau \\ k_2 :: \tau \\ \tau \in \mathcal{T}}
{eq(k_1,k_2) :: Rule}
\end{mathpar}
\caption{Type judgments for Boolean equality rules with probabilistic types}
\label{fig:ptypes}
\end{figure}
}

Taking these type inference rules above, we can run an example inference on the sample training set in Figure~\ref{fig:tset}. The result is the probabilistic type \texttt{bar}$:[bool=3,int=1]$. By setting $M=3, N=1$, we resolve it to \texttt{bar}$::bool$.

A user may implement overlapping inference rules, for example in {\scriptsize INT\_RESOLVE}.
This not a subtyping relation, as the predicates over the counts may not be strictly overlapping.
%TODO this make no sense - address ambiguity more clearly
We simply allow the double colon (::) to type a single keyword in multiple ways.
Because the only typing judgments on a basic type are terminal, {\scriptsize EQ\_RULE}, there is no problem with ambiguity as we would have in a traditional lambda calculus that included function application.

\begin{mathpar}
\hspace{-0.8cm} %no idea why, but this is better than \centering here
\inferrule* [Right=int\_resolve]
{k:\ptype \\ \ptype \models int \geq 1}
{k :: int}
\and
\inferrule* [Right=order\_rule]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}

In the case that there is not enough evidence to resolve a probabilistic type to a basic type, no type-dependent rules may be learned over that keyword.
We are still able to learn rules such as {\scriptsize ORDER}, which do not require any resolved type.

>>>>

Fig.~\ref{fig:ptypes} provides a fragment of a calculus for type judgments for deriving Boolean equality rules with probabilistic types.
When there is no evidence for a type $\tau$ we call this the empty type. 
Initially, all keywords are set to the empty set ({\scriptsize EMPTY}).
In the {\scriptsize BOOL\_RESOLVE} rule, a keyword with a probabilistic type $k : \ptype \in \tilde{\mathcal{T}}$ can be resolved to a basic type when the $\ptype$ satisfies the predicate $p_{bool}$, \ie\ the type has enough evidence.
The definition of enough evidence must be empirically determined by the user depending on the quality of the learning set.

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=Empty]
{\\}
{k:[\ast = 0]}
\and
\inferrule* [Right=bool\_int]
{(k,v) \in C \\ v \in \{0,1\} \\ k:\ptype}
{k:\ptype\update[int\pluseq 1,bool\pluseq 1]}
\and
\inferrule* [Right=bool]
{(k,v) \in C \\ v \in \{ON,OFF\} \\ k:\ptype}
{k:\ptype\update[bool\pluseq 1]}
\\
\and
\inferrule* [Right=int\_resolve]
{k:\ptype \\ p_{int}(\ptype)}
{k :: int}
\and
\inferrule* [Right=bool\_resolve]
{k:\ptype \\ p_{bool}(\ptype)}
{k :: bool}
\\
\and
\inferrule* [Right=eq\_rule]
{k_1 :: \tau \\ k_2 :: \tau \\ \tau \in \mathcal{T}}
{eq(k_1,k_2) :: Rule}
\end{mathpar}
\caption{Type judgments for a probabilistic type system with $\mathcal{T} = \{bool,int\}$ and an equality rule}
\label{fig:ptypes}
\end{figure}
}

In order to define predicates, we use the notation $|\tau_i\ptype|$ to select $c_i$ from a $\ptype \in \tilde{\mathcal{T}}$, 
  and the symbol $*$ as a wildcard selector for all basic types in a $\ptype$.
As an example, we can set the predicate $p_{bool}(\ptype)= |bool\ptype| \geq 3 \land |\ast\ptype| \leq 1$.
Then taking these type inference rules above, we can run an example inference on the sample training set in Figure~\ref{fig:tset}.
The result is the probabilistic type \texttt{bar}$:[bool=3,int=1]$, which is then resolved to \texttt{bar}$::bool$.

A user may pick predicates for probabilistic type resolution that result in overlapping inference rules
For example, if a user picks $p_{int}(\ptype)= |int\ptype| \geq 1$, then {\scriptsize INT\_RESOLVE} overlaps with {\scriptsize BOOL\_RESOLVE}.
This not a not simply a subtyping relation, as the predicates over the counts may not be strictly overlapping.
We simply allow the double colon (::) to type a single keyword in multiple ways.
Because the only typing judgments on a basic type are terminal, {\scriptsize EQ\_RULE}, there is no problem with ambiguity as we would have in a traditional lambda calculus that included function application.

In the case that there is not enough evidence to resolve a probabilistic type to a basic type, no type-dependent rules may be learned over that keyword.
However, we are still able to learn rules such as {\scriptsize ORDER}, which do not require any resolved type.

\begin{mathpar}
\hspace{-0.8cm} %no idea why, but this is better than \centering here
\and
\inferrule* [Right=order\_rule]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}
